%!TEX root = IntensionalSemantics.tex

% Add Iatridou on scope of negative DPs wrt modal contexts

\chapter{DPs and Scope in Modal Contexts}

\chapterprecishere{We discuss ambiguities that arise when DPs occur in modal contexts.}

\minitoc

\section{\emph{Specific} vs. \emph{Non-specific} as a Scope Ambiguity}

When a DP appears inside the clausal or VP complement of a modal predicate\footnote{We will be using the terms ``modal operator'' and ``modal predicate'' in their widest sense here, to include modal auxiliaries (``modals''), modal main verbs and adjectives, attitude predicates, and also modalizing sentence-adverbs like \emph{possibly}. }, there is often a so-called \emph{specific}-\emph{non-specific} ambiguity. A classic example is \Next, which contains the DP \emph{a plumber} inside the infinitive complement of \emph{want}.

\ex. John wants to marry a plumber.

According to the non-specific reading, every possible world in which John gets what he wants is a world in which there is a plumber whom he marries. According to the specific reading, there is a plumber in the actual world whom John marries in every world in which he gets what he wants. We can imagine situations in which one of the readings is true and the other one false.

For example, suppose John thinks that plumbers make ideal spouses, because they can fix things around the house. He has never met one so far, but he definitely wants to marry one. In this scenario, the non-specific reading is true, but the specific reading is false. What all of John's desire-worlds have in common is that they have a plumber getting married to John in them. But it's not the same plumber in all those worlds. In fact, there is no particular individual (actual plumber or other) whom he marries in every one of those worlds.

For a different scenario, suppose that John has fallen in love with Robin and wants to marry Robin. Robin happens to be a plumber, but John doesn't know this; in fact, he wouldn't like it and might even call off the engagement if he found out. Here the specific reading is true, because there is an actual plumber, viz. Robin, who gets married to John in every world in which he gets what he wants. The non-specific reading is false, however, because the worlds which conform to John's wishes actually do not have him marrying a plumber in them. In his favorite worlds, he marries Robin, who is not a plumber in those worlds.

When confronted with this second scenario, you might, with equal justification, say `John wants to marry a plumber', or `John \emph{doesn't} want to marry a plumber'. Each can be taken in a way that makes it a true description of the facts -- although, of course, you cannot assert both in the same breath. This intuition fits well with the idea that we are dealing with a genuine ambiguity.\footnote{In much of the literature the specific/non-specific contrast is also called \emph{de dicto}/\emph{de re}. As of the Spring 2015 edition of these notes, following discussions with Angelika Kratzer, we do not employ that terminology. But one might wonder what was behind the Latin terminology ``de re'' (lit.: `of the thing') and ``de dicto'' (lit.: `of what is said')? Apparently, the term ``de dicto'' is to indicate that on this reading, the \emph{words} which I, the speaker, am using to describe the attitude's content, are the same (at least as far as the relevant DP is concerned) as the words that the subject herself would use to express her attitude. Indeed, if we asked the John in our example what he wants, then in the first scenario he'd say ``marry \emph{a plumber}'', but in the second scenario he would not use these words. The term ``de re'', by contrast, indicates that there is a common \emph{object} (here: Robin) whom I (the speaker) am talking about when I say ``a plumber'' in my report and whom the attitude holder would be referring to if he were to express his attitude in his own words. E.g., in our second scenario, John might say that he wanted to marry ``Robin'', or ``this person here'' (pointing at Robin). He'd thus be referring to the same \emph{person} that I am calling ``a plumber'', but wouldn't use that same description.

Don't take this ``definition'' of the terms too seriously, though! The terminology is much older than any precise truth-conditional analysis of the two readings, and it does not, in hindsight, make complete sense.}

Let's look at another example:

\ex. John believes that your abstract will be accepted.

Here the relevant DP in the complement clause of the verb \emph{believe} is \emph{your abstract}. Again, we detect an ambiguity, which is brought to light by constructing different scenarios.

\begin{enumerate}[(i)]
	
	\item John's belief may be about an abstract that he reviewed, but since the abstract is anonymous, he doesn't know who wrote it. He told me that there was a wonderful abstract about subjacency in Hindi that is sure to be accepted. I know that it was your abstract and inform you of John's opinion by saying \Last. This is the specific reading. In the same situation, the non-specific reading is false: Among John's belief worlds, there are many worlds in which \emph{your abstract will be accepted} is not true or even false. For all he knows, you might have written, for instance, that terrible abstract about Antecedent-Contained Deletion, which he also reviewed and is positive will be rejected.
	
	\item For the other scenario, imagine that you are a famous linguist, and John doesn't have a very high opinion about the fairness of the abstract selection process. He thinks that famous people never get rejected, however the anonymous reviewers judge their submissions. He believes (correctly or incorrectly -- this doesn't matter here) that you submitted a (unique) abstract. He has no specific information or opinion about the abstract's content and quality, but given his general beliefs and his knowledge that you are famous, he nevertheless believes that your abstract will be accepted. This is the non-specific reading. Here it is true in all of John's belief worlds that you submitted a (unique) abstract and it will be accepted. The specific reading of \Last, though, may well be false in this scenario. Suppose -- to flesh it out further -- the abstract you actually submitted is that terrible one about ACD. That one surely doesn't get accepted in every one of John's belief worlds. There may be some where it gets in (unless John is certain it can't be by anyone famous, he has to allow at least the possibility that it will get in despite its low quality). But there are definitely also belief-worlds of his in which it doesn't get accepted.
	
	We have taken care here to construct scenarios that make one of the readings true and the other false. This establishes the existence of two distinct readings. We should note, however, that there are also many possible and natural scenarios that simultaneously support the truth of \emph{both} readings. Consider, for instance, the following third scenario for sentence \Last.
	
	\item John is your adviser and is fully convinced that your abstract will be accepted, since he knows it and in fact helped you when you were writing it. This is the sort of situation in which both the non-specific and the specific reading are true. It is true, on the one hand, that the sentence \emph{your abstract will be accepted} is true in every one of John's belief worlds (non-specific reading). And on the other hand, if we ask whether the abstract which you actually wrote will get accepted in each of John's belief worlds, that is likewise true (specific reading).
	
	In fact, this kind of ``doubly verifying'' scenario is very common when we look at actual uses of attitude sentences in ordinary conversation. There may even be many cases where communication proceeds smoothly without either the speaker or the hearer making up their minds as to which of the two readings they intend or understand. It doesn't matter, since the possible circumstances in which their truth-values would differ are unlikely and ignorable anyway. Still, we \emph{can} conjure up scenarios in which the two readings come apart, and our intuitions about those scenarios do support the existence of a semantic ambiguity.
\end{enumerate}

In the paraphrases by which we have elucidated the two readings of our examples, we have already given away the essential idea of the analysis that we will adopt: We will treat non-specific/specific ambiguities as ambiguities of \emph{scope}. The non-specific readings, it turns out, are the ones which we predict without further ado if we assume that the position of the DP at LF is within the modal predicate's complement. (That is, it is either \emph{in situ} or QRed within the complement clause.) For example:

\exi. \label{first} John wants [ [ a plumber]$_1$ [ \textsc{pro}$_2$ to marry t$_1$]]

\exi. John believes [ the abstract-by-you will-be-accepted]

To obtain the specific readings, we apparently have to QR the DP to a position above the modal predicate, minimally the VP headed by \emph{want} or \emph{believe}.

\exi. [ a plumber]$_1$ [ John wants [ \textsc{pro}$_2$ to marry t$_1$]]

\exi. [ the abstract-by-you]$_1$ [ John believes t$_1$ will-be-accepted]]

\begin{exercise}
	
	Calculate the interpretations of the four structures in \ref{first}--\Last, and determine their predicted truth-values in each of the (types of) possible worlds that we described above in our introduction to the ambiguity.
	
	Some assumptions to make the job easier: (i) Assume that \ref{first} and \LLast are evaluated with respect to a variable assignment that assigns John to the number 2. This assumption takes the place of a worked out theory of how controlled PRO is interpreted. (ii) Assume that \emph{abstract-by-you} is an unanalyzed one-place predicate. This takes the place of a worked out theory of how genitives with a non-possessive meaning are to be analyzed. \eex
\end{exercise}

\section{Raised subjects}

In the examples of specific-non-specific ambiguities that we have looked at so far, the surface position of the DP in question was inside the modal predicate's clausal or VP-complement. We saw that if it stays there at LF, a non-specific reading results, and if it covertly moves up above the modal operator, we get a specific reading. In the present section, we will look at cases in which a DP that is superficially \emph{higher} than a modal operator can still be read non-specific. In these cases, it is the specific reading which we obtain if the LF looks essentially like the surface structure, and the non-specific reading for which we apparently have to posit a non-trivial covert derivation.

\subsection{Examples of non-specific readings for raised subjects}

Suppose I come to my office one morning and find the papers and books on my desk in different locations than I remember leaving them the night before. I say:

\ex. \label{some} Somebody must have been here (since last night).

On the assumptions we have been making, \emph{somebody} is base-generated as the subject of the VP \emph{be here} and then moved to its surface position above the modal. So \Last has the following S-structure, which is also an interpretable LF.

\exi. \label{dere} somebody [ $\lambda_2$ [ [ must $R$] [ t$_2$ have-been-here]]]

What does \Last mean? The appropriate reading for \emph{must} here is epistemic, so suppose the variable $R$ is mapped to the relation $\bigl[\lambda w.\lambda w'.\ w'$ is compatible with what I believe in $w\bigr]$. Let $w_{0}$ be the utterance world. Then the truth-condtion calculated by our rules is as follows.

\ex. $\exists x [x$ is a person in $w_{0}$ \& \\
$\forall w'[w'$ is compatible with what I believe in $w_{0}$ $\rightarrow\ x$ was here in $w'$]]

But this is not the intended meaning. For \Last to be true, there has to be a person who in every world compatible with what I believe was in my office. In other words, all my belief-worlds have to have one and the same person coming to my office. But this is not what you intuitively understood me to be saying about my belief-state when I said \ref{some}. The context we described suggests that I do not know (or have any opinion about) which person it was that was in my office. For all I know, it might have been John, or it might have been Mary, or it have been this stranger here, or that stranger there. In each of my belief-worlds, somebody or other was in my office, but no one person was there in all of them. I do not believe of anyone in particular that he or she was there, and you did not understand me to be saying so when I uttered \ref{some}. What you did understand me to be claiming, apparently, was not \Last but \Next.

\ex. \label{dedic} $\forall w' [ w'$ is compatible with what I believe in $w_{0}$\\
\null\hfill$\rightarrow\ \exists x\ [x$ is a person in $w'$ \& $x$ was here in $w']]$

In other words -- to use the terminology we introduced in the last section -- the DP \emph{somebody} in \ref{some} appears to have a non-specific reading.

How can sentence \ref{some} have the meaning in \Last? The LF in \ref{dere}, as we saw, means something else; it expresses a specific reading, which typically is false when \ref{some} is uttered sincerely. So there must be another LF. What does it look like and how is it derived? One way to capture the intended reading, it seems, would be to generate an LF that's essentially the same as the underlying structure we posited for \ref{some}, i.e., the structure \emph{before} the subject has raised:

\ex. [IP e [I$'$ [ must $R$] [ somebody have-been-here]]]

\Last means precisely \LLast (assuming that the unfilled Spec-of-IP position is semantically vacuous), as you can verify by calculating its interpretation by our rules. So is \Last (one of) the LF(s) for \ref{some}, and what assumption about syntax allow it to be generated? Or are there other -- perhaps less obvious, but easier to generate -- candidates for the non-specific LF-structure of \ref{some}?

Before we get into these question, let's look at a few more examples. Each of the following sentences, we claim, has a non-specific reading for the subject, as given in the accompanying formula. The modal operators in the examples are of a variety of syntactic types, including modal auxiliaries, main verbs, adjectives, and adverbs.

\ex. \label{everymay} Everyone in the class may have received an A.\\
$\exists w'[w'$ conforms to what I believe in $w$ \&\\
\null\hfill$\forall x[x$ is in this class in $w'\ \rightarrow\ x$ received an A in $w'$]].

\ex. At least two semanticists have to be invited.\\
$\forall w'[w'$ conforms to what is desirable in $w$\\
\null\hfill$ \rightarrow\ \exists_2 x$\ [$x$ is a semanticist in $w'$ \& $x$ is invited in $w'$]].

\ex. \label{ny} Somebody from New York is expected to win the lottery.\\
$\forall w'[w'$ conforms to what is expected in $w$\\
\null\hfill$ \rightarrow\ \exists x[x$ is a person from NY in $w'$ \& $x$ wins the lottery in $w'$]]

\ex. \label{nyli} Somebody from New York is likely to win the lottery.\\
$\forall w'[w'$ is as likely as any other world, given I know in $w$\\
\null\hfill$ \rightarrow\ \exists x[x$ is a person from NY in $w'$ \& $x$ wins the lottery in $w'$]]\footnote{Hopefully the exact analysis of the modal operators \emph{likely} and \emph{probably} is not too crucial for the present discussion, but you may still be wondering about it. As you see in our formula, we are thinking of \emph{likely} (\emph{probably}) as a kind of epistemic necessity operator, i.e., a universal quantifier over a set of worlds that is somehow determined by the speaker's knowledge. (We are focussing on the ``subjective probability'' sense of these words. Perhaps there is a also an ``objective probability'' reading that is circumstantial rather than epistemic.) What is the difference then between \emph{likely} and e.g. epistemic \emph{must} (or \emph{necessary} or \emph{I believe that})? Intuitively, `it is likely that p' makes a weaker claim than `it must be the case that p'. If both are universal quantifiers, then, it appears that \emph{likely} is quantifying over a smaller set than \emph{must}, i.e., over only a proper subset of the worlds that are compatible with what I believe. The difference concerns those worlds that I cannot strictly rule out but regard as remote possibilities. These worlds are included in the domain for \emph{must}, but not in the one for \emph{likely}. For example, if there was a race between John and Mary, and I am willing to bet that Mary won but am not completely sure she did, then those worlds where John won are remote possibilities for me. They are included in the domain of \emph{must}, and so I will not say that Mary \emph{must} have won, but they are not in the domain quantified over by \emph{likely}, so I do say that Mary is \emph{likely} to have won.\\
\indent This is only a very crude approximation, of course. For one thing, probability is a gradable notion. Some things are more probable than others, and where we draw the line between what's probable and what isn't is a vague or context-dependent matter. Even \emph{must}, \emph{necessary} etc. arguably don't really express complete certainty (because in practice there is hardly anything we are completely certain of), but rather just a very high degree of probability. For more discussion of \emph{likely}, \emph{necessary}, and other graded modal concepts in a possible worlds semantics, see e.g. Kratzer 1981, 1991.\\
\indent A different approach may be that \emph{likely} quantifies over the same set of worlds as \emph{must}, but with a weaker, less than universal, quantificational force. I.e., `it is likely that p' means something like p is true in \emph{most} of the worlds conforming to what I know. A \emph{prima facie} problem with this idea is that presumably every proposition is true in infinitely many possible worlds, so how can we make sense of cardinal notions like `more' and `most' here? But perhaps this can be worked out somehow. }

\ex. One of these two people is probably infected.\\
$\forall w'[w'$ is as likely as any other world, given what I know in $w$\\
\null\hfill$ \rightarrow\ \exists x[x$ is one of these two people \& $x$ is in infected in $w'$]]

To bring out the intended non-specific reading of the last example (to pick just one) imagine this scenario: We are tracking a dangerous virus infection and have sampled blood from two particular patients. Unfortunately, we were sloppy and the blood samples ended up all mixed up in one container. The virus count is high enough to make it quite probable that one of the patients is infected but because of the mix-up we have no evidence about which one of them it may be. In this scenario, \Last appears to be true. It would not be true under a specific reading, because neither one of the two people is infected in every one of the likely worlds.

A word of clarification about our empirical claim: We have been concentrating on the observation that non-specific readings are \emph{available}, but have not addressed the question whether they are the \emph{only} available readings or coexist with equally possible specific readings. Indeed, some of the sentences in our list appear to be ambiguous: For example, it seems that \ref{ny} could also be understood to claim that there is a particular New Yorker who is likely to win (e.g., because he has bribed everybody). Others arguably are not ambiguous and can only be read non-specific. This is what \citet{fintel-iatridou:2003:ec} claim about sentences like \ref{everymay}. They note that if \ref{everymay} also allowed a specific reading, it should be possible to make coherent sense of \Next.

\ex. Everyone in the class may have received an A. But not everybody did.

In fact, \Last sounds contradictory, which they show is explained if only the non-specific reading is permitted by the grammar. They conjecture that this is a systematic property of epistemic modal operators (as opposed to deontic and other types of modalities). Epistemic operators always have widest scope in their sentence.

So there are really two challenges here for our current theory. We need to account for the existence of non-specific readings, and also for the absence, in at least some of our examples, of specific readings. We will be concerned here exclusively with the first challenge and will set the second aside. We will aim, in effect, to set up the system so that all sentences of this type are in principle ambiguous, hoping that additional constraints that we are not investigating here will kick in to exclude the specific readings where they are missing.

To complicate the empirical picture further, there are also examples where raised subjects are unambiguously specific. Such cases have been around in the syntactic literature for a while, and they have received renewed attention in the work of Lasnik and others. To illustrate just one of the systematic restrictions, negative quantifiers like \emph{nobody} seem to permit only surface scope (i.e., wide scope) with respect to a modal verb or adjective they have raised over.

\ex. Nobody from New York is likely to win the lottery.

\Last does not have a non-specific reading parallel to the one for \ref{nyli} above, i.e., it cannot mean that it is likely that nobody from NY will win. It can only mean that there is nobody from NY who is likely to win. This too is an issue that we set aside.

In the next couple of sections, all that we are trying to do is find and justify a mechanism by which the grammar is capable to generate both specific and non-specific readings for subjects that have raised over modal operators. It is quite conceivable, of course, that the nature of the additional constraints which often exclude one reading or the other is ultimately relevant to this discussion and that a better understanding of them may undermine our conclusions. But this is something we must leave for further research.

\subsection{Syntactic ``Reconstruction''}

Given that the non-specific reading of \ref{some} we are aiming to generate is equivalent to the formula in \ref{dedic}, an obvious idea is that there is an LF which is essentially the pre-movement structure of this sentence, i.e., the structure prior to the raising of the subject above the operator. There are a number of ways to make such an LF available.

One option, most defended in \citet{sauerland-elbourne:2002:total}, is to assume that the raising of the subject can happen in a part of the derivation which only feeds PF, not LF. In that case, the subject simply stays in its underlying VP-internal position throughout the derivation from DS to LF. (Recall that quantifiers are interpretable there, as they generally are in subject positions.)

Another option is a version of the so-called Copy Theory of movement introduced in \citet{chomsky:1993:minimalist}. This assumes that movement generally proceeds in two separate steps, rather than as a single complex operation as we have assumed so far. Recall that in H\amp K, it was stipulated that every movement effects the following four changes:
\begin{enumerate}
	[(i)] 
	\item a phrase $\alpha$ is deleted, 
	\item an index \emph{i} is attached to the resulting empty node (making it a so-called trace, which the semantic rule for ``Pronouns and Traces'' recognizes as a variable), 
	\item a new copy of $\alpha$ is created somewhere else in the tree (at the ``landing site''), and 
	\item the sister-constituent of this new copy gets another instance of the index \emph{i} adjoined to it (which the semantic rule of Predicate Abstraction recognizes as a binder index). 
\end{enumerate}

If we adopt the Copy Theory, we assume instead that there are three distinct operations:
\begin{description}
	
	\item[``Copy'':] Create a new copy of $\alpha$ somewhere in the tree, attach an index \emph{i} to the original $\alpha$ , and adjoin another instance of \emph{i} to the sister of the new copy of $\alpha$ . (= steps (ii), (iii), and (iv) above)
	
	\item[``Delete Lower Copy'':] Delete the original $\alpha$ . (= step (i) above)
	
	\item[``Delete Upper Copy'':] Delete the new copy of $\alpha$ and both instances of \emph{i}.
\end{description}

\noindent The Copy operation is part of every movement operation, and can happen anywhere in the syntactic derivation. The Delete operations happen at the end of the LF derivation and at the end of the PF deletion. We have a choice of applying either Delete Lower Copy or Delete Upper Copy to each pair of copies, and we can make this choice independently at LF and at PF. (E.g., we can do Copy in the common part of the derivation and than Delete Lower Copy at LF and Delete Upper Copy at PF.) If we always choose Delete Lower Copy at LF, this system generates exactly the same structures and interpretations as the one from H\amp K. But if we exercise the Delete Upper Copy option at LF, we are effectively undoing previous movements, and this gives us LFs with potentially new interpretations. In the application we are interested in here, we would apply the Copy step of subject raising before the derivation branches, and then choose Delete Lower Copy at PF but Delete Upper Copy at LF. The LF will thus look as if the raising never happened, and it will straightforwardly get the desired non-specific reading.

If the choice between the two Delete operations is generally optional, we in principle predict ambiguity wherever there has been movement. Notice, however, first, that the two structures will often be truth-conditionally equivalent (e.g. when the moved phrase is a name), and second, that they will not always be both interpretable. (E.g., if we chose Delete Upper Copy after QRing a quantifier from object position, we'd get an uninterpretable structure, and so this option is automatically ruled out.) Even so, we predict lots of ambiguity. Specifically, since raised subjects are always interpretable in both their underlying and raised locations, we predict all raising structures where a quantificational DP has raised over a modal operator (or over negation or a temporal operator) to be ambiguous. As we have already mentioned, this is not factually correct, and so there must be various further constraints that somehow restrict the choices. (Similar comments apply, of course, to the option we mentioned first, of applying raising only on the PF-branch.)

Yet another solution was first proposed by \citet{may:1977:thesis}: May assumed that QR could in principle apply in a ``downward'' fashion, i.e., it could adjoin the moved phrase to a node that doesn't contain its trace. Exercising this option with a raised subject would let us produce the following structure, where the subject has first raised over the modal and then QRed below it.

\ex. t$_j$ $\lambda_i$ [ must-R [ someone $\lambda_j$ [ t$_i$ have been here]]]

As it stands, this structure contains at least one free variable (the trace \emph{t}$_j$) and can therefore not possibly represent any actual reading of this sentence. May further assumes that traces can in principle be deleted, when their presence is not required for interpretability. This is not yet quite enough, though to make \Last interpretable, at least not within our framework of assumptions, for \Next is still not a candidate for an actual reading of \ref{some}.

\ex. $\lambda_i$ [ must-R [ someone $\lambda_j$ [ t$_i$ have been here]]]

We would need to assume further that the topmost binder index could be deleted along with the unbound trace, and also that the indices \emph{i} and \emph{j} can be the same, so that the raising trace \emph{t}$_j$ is bound by the binding-index created by QR. If these things can be properly worked out somehow, then this is another way to generate the non-specific reading. Notice that the LF is not exactly the same as on the previous two approaches, since the subject ends up in an adjoined position rather than in its original argument position, but this difference is obviously without semantic import.

What all of these approaches have in common is that they place the burden of generating the non-specific reading for raised subjects on the syntactic derivation. Somehow or other, they all wind up with structures in which the subject is lower than it is on the surface and thereby falls within the scope of the modal operator. They also have in common that they take the modal operator (here the auxiliary, in other cases a main predicate or an adverb) to be staying put. I.e., they assume that the non-specific readings are not due to the modal operator being covertly higher than it seems to be, but to the subject being lower. Approaches with these features will be said to appeal to ``syntactic reconstruction'' of the subject.\footnote{This is a very broad notion of ``reconstruction'', where basically any mechanism which puts a phrase at LF in a location nearer to its underlying site than its surface site is called ``reconstruction''. In some of the literature, the term is used more narrowly. For example, May's downward QR is sometimes explicitly contrasted with genuine reconstruction, since it places the quantifier somewhere else than exactly where it has moved from. }

\subsection{Some Alternatives to Syntactic Reconstruction}\label{sem}

Besides (some version of) syntactic reconstruction, there are many other ways in which one try to generate non-specific readings for raised subjects. Here are some other possibilities that have been suggested and/or readily come to mind. We will see that some of them yield exactly the non-specific reading as we have been describing it so far, whereas others yield a reading that is very similar but not quite the same. We will confine ourselves to analyses which involve no or only minor changes to our system of syntactic and semantic assumptions. Obviously, if departed from these further, there would be even more different options, but even so, there seem to be quite a few.

\subsubsection{1. Raising the modal operator, variant 1: no trace}

Conceivably, an LF for the non-specific reading of \ref{some} might be derived from the S-structure (=\ref{dere}) by covertly moving \emph{must} (and its covert \emph{R}-argument) up above the subject. This would have to be a movement which leaves no (semantically non-vacuous) trace. Given our inventory of composition rules, the only type that the trace could have to make the structure containing it interpretable would be the type of the moved operator itself (i.e. \angles{st,t}). If it had that type, however, the movement would be semantically inconsequential, i.e., the structure would mean exactly the same as \ref{dere}. So this would not be a way to provide an LF for the non-specific reading. If there was no trace left however (and also no binder index introduced), we indeed would obtain the non-specific reading.
\begin{exercise}
	Prove the claims we just made in the previous paragraph. Why is no type for the trace other than \angles{st,t} possible? Why is the movement semantically inert when this type is chosen? How does the correct intended meaning arise if there is no trace and binder index? \eex 
\end{exercise}

\subsubsection{2. Raising the modal operator, variant 2: trace of type s}

[Requires slightly modified inventory of composition rules. Derives an
interpretation that is not quite the same as the non-specific
reading we have assumed so far. Rather, it is a ``narrow-Q, R-de-re''
interpretation in the sense discussed in the next chapter.]

\subsubsection{3. Higher type for trace of raising, variant 1: type \angles{et,t} }

[Before reading this section, read and do the exercise on p.212/3 in H\amp K]

\absatz So far in our discussion, we have taken for granted that the LF which corresponds to the surface structure, viz. \ref{dere}, gives us the specific reading. This, however, is correct only on the tacit assumption that the trace of raising is a variable of type e. If it is part of our general theory that all variables, or at least all interpretable binder indices (hence all bound variables), in our LFs are of type e, then there is nothing more here to say. But it is not \emph{prima facie} obvious that we must or should make this general assumption, and if we don't, then the tree in \ref{dere} is not really one single LF, but the common structure for many different ones, which differ in the type chosen for the trace. Most of the infinitely many semantic types we might assign to this trace will lead to uninterpretable structures, but there turns out to be one other choice besides e that works, namely \angles{et,t}:

\exi. somebody $\lambda_{2,\angles{et,t}}$ [ [ must $R$] [ $t_{2,\angles{et,t}}$ have-been-here]]

\Last is interpretable in our system, but again, as above, the predicted interpretation is not exactly the non-specific reading as we have been describing it so far, but a ``narrow-Q, R-de-re'' reading.
\begin{exercise}
	
	Using higher-type traces to ``reverse'' syntactic scope-relation is a trick which can be used quite generally. It is useful to look at a non-intensional example as a first illustration. \Next contains a universal quantifier and a negation, and it is scopally ambiguous between the readings in (a) and (b).
	
	\ex. Everything that glitters is not gold. \a. $\forall x[ x$ glitters $\rightarrow\ \ensuremath{\neg} x$ is gold] \hfill``surface scope'' \b. $\ensuremath{\neg}\forall x[ x$ glitters $\rightarrow\ x$ is gold] \hfill``inverse scope''
	
	We could derive the inverse scope reading for \Last by generating an LF (e.g. by some version of syntactic reconstruction") in which the \emph{every}-DP is below \emph{not}. Interestingly, however, we can also derive this reading if the \emph{every}-DP is in its raised position above \emph{not} but its trace has the type \angles{\angles{e,t},t}.
	
	Spell out this analysis. (I.e., draw the LF and show how the inverse-scope interpretation is calculated by our semantic rules.) \eex 
\end{exercise}
\begin{exercise}
	
	Convince yourself that there are no other types for the raising trace besides e and \angles{et,t} that would make the structure in \ref{dere} interpretable. (At least not if we stick exactly to our current composition rules.) \eex 
\end{exercise}

\subsubsection{4. Higher type for trace of raising, variant 2: type \angles{s,\angles{et,t}}}

If we want to get \emph{exactly} the non-specific reading that results from syntactic reconstruction out of a surface-like LF of the form \ref{dere}, we must use an even higher type for the raising trace, namely \angles{s,\angles{\angles{e,t},t}}, the type of the intension of a quantifier. As you just proved in the exercise, this is not possible if we stick to exactly the composition rules that we have currently available. The problem is in the VP: the trace in subject position is of type \angles{s,\angles{\angles{e,t},t}} and its sister is of type \angles{e,t}. These two connot combine by either FA or IFA, but it works if we employ another variant of functional application.\footnote{Notice that the problem here is kind of the mirror image of the problem that led to the introduction of ``Intensional Functional Application'' in H\amp K, ch. 12. There, we had a function looking for an argument of type \angles{s,t}, but the sister node had an extension of type t. IFA allowed us to, in effect, construct an argument with an added ``s'' in its type. This time around, we have to get rid of an ``s'' rather than adding one; and this is what EFA accomplishes. \\
\indent So we now have three different ``functional application''-type rules altogether in our system: ordinary FA simply applies $\sv{\beta}^w$ to $\sv{\gamma}^w$; IFA applies $\sv{\beta}^w$ to $\lambda w'.\sv{\gamma}^{w'}$; and EFA applies $\sv{\beta}^w(w)$ to $\sv{\gamma}^w$. At most one of them will be applicable to each given branching node, depending on the type of $\sv{\gamma}^w$.\\
\indent Think about the situation. Might there be other variant functional application rules?}

\ex. \emph{Extensionalizing Functional Application} (EFA)\\
If $\alpha$ is a branching node and $\{\beta,\gamma\}$ the set of its daughters, then, for any world $w$ and assignment $g$: \\
if $\sv{\beta}^{w,g}(w)$ is a function whose domain contains $\sv{\gamma}^{w,g}$,\\
then $\sv{\alpha}^{w,g} = \sv{\beta}^{w,g}(w)(\sv{\gamma}^{w,g})$.

\begin{exercise}
	
	Calculate the truth-conditions of \ref{dere} under the assumption that the trace of the subject quantifier is of type \angles{s,\angles{\angles{e,t},t}}. \eex 
\end{exercise}

\subsubsection{Can we choose between all these options?}\label{semrec}

Two of the methods we tried derived readings in which the raised subject's \emph{quantificational determiner} took scope below the world-quantifier in the modal operator, but the raised subject's \emph{restricting NP} still was evaluated in the utterance world (or the evaluation world for the larger sentence, whichever that may be). It is difficult to assess whether these readings are actually available for the sentences under consideration, and we will postpone this question to a later section. We would like to argue here, however, that even if these readings are available, they cannot be the \emph{only} readings that are available for raised subjects besides their wide-scope readings. In other words, even if we allowed one of the mechanisms that generated these sort of hybrid readings, we would still need another mechanism that gives us, for at least some examples, the ``real'' non-specific readings that we obtain e.g. by syntactic reconstruction. The relevant examples that show this most clearly involve DPs with more descriptive content than \emph{somebody} and whose NPs express clearly contingent properties.

\ex. A neat-freak must have been here.

\enlargethispage{36pt}If I say this instead of our original \ref{some} when I come to my office in the morning and interpret the clues on my desk, I am saying that every world compatible with my beliefs is such that someone who is a neat-freak \emph{in that world} was here in that world. Suppose there is a guy, Bill, whom I know slightly but not well enough to have an opinion on whether or not he is neat. He may or not be, for all I know. So there are worlds among my belief worlds where he is a neat-freak and worlds where he is not. I also don't have an opinion on whether he was or wasn't the one who came into my office last night. He did in some of my belief worlds and he didn't in others. I am implying with \Last, however, that if Bill isn't a neat-freak, then it wasn't him in my office. I.e., \Last \emph{is} telling you that, even if I have belief-worlds in which Bill is a slob and I have belief-worlds in which (only) he was in my office, I do not have any belief-worlds in which Bill is a slob \emph{and} the only person who was in my office. This is correctly predicted if \Last expresses the ``genuine'' non-specific reading in \Next, but not if it expresses the ``hybrid'' reading in \NNext.

\ex. $\forall w'[w'$ is compatible with what I believe in $w_{0}\ \rightarrow$ \\
\null\hfill$\exists x[x$ is a neatfreak \emph{in $w'$} and $x$ was here in $w'$]]

\ex. $\forall w'[w'$ is compatible with what I believe in $w_{0}\ \rightarrow$ \\
\null\hfill$\exists x[x$ is a neatfreak \emph{in $w_{0}$} and $x$ was here in $w'$]]

We therefore conclude the mechanisms 2 and 3 considered above (whatever there merits otherwise) cannot supplant syntactic reconstruction or some other mechanism that yields readings like \LLast.

This leaves only the first and fourth options that we looked at as potential competitors to syntactic reconstruction, and we will focus the rest of the discussion on how we might be able to tease apart the predictions that these mechanisms imply from the ones of a syntactic reconstruction approach.

As for moving the modal operator, there are no direct bad predictions that we are aware of with this. But it leads us to expect that we might find not only scope ambiguities involving a modal operator and a DP, but also scope ambiguities between two modal operators, since one of them might covertly move over the other. It seems that this never happens. Sentences with stacked modal verbs seem to be unambiguous and show only those readings where the scopes of the operators reflect their surface hierarchy.

\ex. \a. I have to be allowed to graduate. \b. \#I am allowed to have to graduate. 

Of course, this might be explained by appropriate constraints on the movement of modal operators, and such constraints may even come for free in a the right synatctic theory. Also, we should have a much more comprehensive investigation of the empirical facts before we reach any verdict. If it is true, however, that modal operators only engage in scope interaction with DPs and never with each other, then a theory which does not allow any movement of modals at all could claim the advantage of having a simple and principled explanation for this fact.

What about the ``semantic reconstruction'' option, where raised subjects can leave traces of type \angles{s,\angles{et,t}} and thus get narrow scope semantically without ending up low syntactically? This type of approach has been explored quite thoroughly and defended with great sophistication. We can only sketch the main objections to it here and must leave it to the reader to consult the literature for an informed opinion.

\subsubsection{Scope reconstruction and Condition C}

An  %\mycite{lebeaux:1998}\mycite{heycock:1995}\mycite{romero:1997}\mycite{fox:1999}\mycite{fox:2000}\mycite{sternefeld:1997}\mycite{sharvit:1998}
example from Fox (1999) (building on Lebeaux 1994 and Heycock 1995):

\ex. \a. A student of his$_{1}$ seems to David$_1$ to be at the party.\\
\null\hfill $^{OK}$specific, $^{OK}$non-specific 
\b. A student of David's$_{1}$ seems to him$_{1}$ to be at the party.\\
\null\hfill $^{OK}$specific, {\small *}non-specific

Sketch of argument: If Cond. C is formulated in terms of c-command relations and applies at LF, it will distinguish between specific and non-specific readings only if those involve LFs with different hierarchical relations.

\enlargethispage{36pt}

\subsubsection{Raising of idiom chunks}

\ex. The cat seems to be out of the bag.

\ex. ?Advantage might have been taken of them.

Sketch of argument: If idioms must be constituents at LF in order to receive their idiomatic interpretations, these cases call for syntactic reconstruction. An additional mechanism of semantic reconstruction via high-type traces is then at best redundant.

\medskip\noindent Tentative conclusion: Syntactic reconstruction (some version of it) provides the best account of non-specific readings for raised subjects.